\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-12-23 adaptive-estimation.tex
Finished writing major proofs and fixing references. Still needs a once through
corrected various typos

%% Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
%% \begin{equation*}
%% n\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{1}{4f^2(0)}+O(n^{-1}),
%% \end{equation*}
%% and 
%% \begin{equation*}
%% \ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
%% \end{equation*}
It is possible to extend Theorem~\ref{thm:adpative_lower_bound} to any other
loss functions using a more subtle version of the Van Trees
inequality~\cite{efroimovich1980information}; see also
\cite{DBLP:journals/corr/abs-1902-08582}.

@@

The starting point for our first estimator is to note that the median of a
distribution minimizes $\E[|X - \theta|]$ over $\theta \in \R$, and
moreover, we have the familiar result (cf.~\cite{VanDerVaart98}, Ch.~21)
that given a sample $X_1^n \simiid P$, if $\theta = \mbox{med}(P)$ and
$P$ has continuous density $f(\cdot - \theta)$ near $\theta$, then
\begin{equation*}
  \sqrt{n}(\mbox{med}(X_1^n) - \theta)
  \cd \normal\left(0, \frac{1}{4 f(0)^2}\right),
\end{equation*}
which is precisely the variance lower bound in
Theorem~\ref{thm:adpative_lower_bound}.  Thus, it it is natural consider a
stochastic gradient procedure for minimizing $\E[|X - \theta|]$. To that end,
let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive
sequence of stepsizes,
and define the sequence
\begin{equation}
  \label{eq:sgd_alg}
  \theta_n = \theta_{n-1} + \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation*}
  B_n = \sgn (X_n - \theta_{n-1}).
\end{equation*}
We make one of two assumptions on the stepsizes $\gamma_n$, which
are relatively standard: we always have $\gamma_n$ non-increasing, and
\begin{subequations}
  For some $0 < \lambda \le 1$,
  \begin{align}
    \label{eqn:lazy-gamma}
    \frac{\gamma_n - \gamma_{n+1}}{\gamma_n^2}
    \to 0, & ~~~
    \sum_n \frac{\gamma_n^\frac{1 + \lambda}{2}}{\sqrt{n}} < \infty
    ~~ \mbox{or} \\
    \gamma_n = o(n^{-2/3}),
    & ~~~
    \sum_n \gamma_n = \infty.
    \label{eqn:stringent-gamma}
  \end{align}
\end{subequations}

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(0)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    Let $\{P_\theta\}_{\theta \in \R}$ be the family of distributions
    with density $f(\cdot - \theta)$, where $f$ has median 0.
    Let $h_n \to h \in \R$, and define the distributions
    $P_n = P_{\theta + h_n/\sqrt{n}}^n$. Then
    \begin{equation*}
      \sqrt{n}\left(\bar{\theta}_n - \theta - h_n / \sqrt{n}\right)
      \mathop{\cd}_{P_n}
      \normal\left(0, \frac{1}{4 f(0)^2}\right),
    \end{equation*}
    and for any bounded, symmetric, and quasi-convex function $L$,
    \begin{align} 
      & \sup_{c < \infty} \limsup_{n \to \infty}
      \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}
      \E_\tau \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[L(W/ 2 f(0)) \right],
        \label{eq:attaining_LAM}
    \end{align}
    where $W \sim \normal(0,1)$. 
    %
    %
  \item \label{item:sgd-ms-convergence} Assume the stepsizes $\gamma_n$
    satisfy both conditions~\eqref{eqn:lazy-gamma}
    and~\eqref{eqn:stringent-gamma}. Let
    $\pi$ be a distribution on $\R$.
    \begin{equation*}
      \int \E\Big[( \bar{\theta}_n - \theta )^2\Big] d\pi(\theta)
      = \frac{1}{4 n f(0)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}

@@

%% Theorem~\ref{thm:sgd} implies that the estimator ${\theta}_n$ defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} attains the maximal ARE, as established by Theorem~\ref{thm:adpative_lower_bound}. %Furthermore, it is locally minimax in the sense that it attains the minimal MSE for all local alternatives of $\theta$. 
%% %
%% The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. In particular, parts (i) and (ii) of Theorem~\ref{thm:sgd} provides conditions under which $\bar{\theta}$ is asymptotically normal and local asymptotically minimax, in the sense that it attains the local asymptotic minimax bound \cite{van2000asymptotic}. Part (iii) provides an additional condition under which $\bar{\theta}$ also converges in its second moment.

@@

  Now let
  $E_n = \E[B_i \mid T_n] = P(X_i \le T_n)$,
  so that $\var(B_i \mid T_n) = E_n(1 - E_n)$. Define also the random
  variable
  \begin{equation*}
    Y_n \defeq \sqrt{n}
    \frac{1}{\sqrt{E_n (1 - E_n)}}
    \bigg[\frac{1}{n}\sum_{i = 1}^n B_i - E_n\bigg],
  \end{equation*}
  and let $F_n(\cdot \mid T_n)$ be its cumulative distribution function.
  Then because
  \begin{equation*}
    \E\left[|B_i - E_n|^3 \mid T_n\right] \le E_n(1 - E_n)
    ~~ \mbox{we have} ~~
    \E\left[\frac{|B_i - E_n|^3}{(E_n(1 - E_n))^{3/2}} \mid T_n\right]
    \le \frac{1}{\sqrt{E_n(1 - E_n)}},
  \end{equation*}
  the Berry-Esseen theorem implies that there exists a constant
  $C \le 1$ such that
  \begin{equation*}
    \sup_t \left|F_n(t \mid T_n) - \Phi(t) \right|
    \le \frac{C}{\sqrt{E_n (1 - E_n)} \sqrt{n}} \wedge 2,
  \end{equation*}
  where $\Phi$ is the standard Gaussian CDF.
  As $E_n (1 - E_n) \cas \frac{1}{4}$ by definition of the median,
  we have
  that (with probability 1)
  \begin{equation*}
    \sup_t \left|F_n(t \mid T_n) - \Phi(t)\right| \le \frac{C}{\sqrt{n}}
    ~~ \mbox{eventually}.
  \end{equation*}
  By dominated convergence and Jensen's inequality we thus obtain
  \begin{equation*}
    \sup_t \left|\mathbb{P}(Y_n \le t) - \Phi(t)\right|
    \le \E\left[
      \sup_t \left|F_n(t \mid T_n) - \Phi(t)\right| \right]
    \to 0,
  \end{equation*}
  which gives that $Y_n \cd \normal(0, 1)$.

>>>>

%% Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
%% \begin{equation*}
%% n\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{1}{4f^2(0)}+O(n^{-1}),
%% \end{equation*}
%% and 
%% \begin{equation*}
%% \ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
%% \end{equation*}
It is possible to extend Theorem~\ref{thm:adpative_lower_bound} to other
loss functions using a more subtle version of the Van Trees
inequality~\cite{efroimovich1980information}; see also
\cite{DBLP:journals/corr/abs-1902-08582}.

@@

The starting point for our first estimator is to note that the median of a
distribution minimizes $\E[|X - \theta|]$ over $\theta \in \R$, and
moreover, we have the familiar result (cf.~\cite{VanDerVaart98}, Ch.~21)
that given a sample $X_1,\ldots,X_n \simiid P$, if $\theta = \mbox{med}(P)$ and
$P$ has continuous density $f(\cdot - \theta)$ near $\theta$, then
\begin{equation*}
  \sqrt{n}(\mbox{med}(X_1^n) - \theta)
  \cd \normal\left(0, \frac{1}{4 f(0)^2}\right),
\end{equation*}
which is precisely the variance lower bound in
Theorem~\ref{thm:adpative_lower_bound}.  Thus, it is natural to consider a
stochastic gradient procedure for minimizing $\E[|X - \theta|]$. To that end,
let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive
sequence of stepsizes,
and define the sequence
\begin{equation}
  \label{eq:sgd_alg}
  \theta_n = \theta_{n-1} + \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation*}
  B_n = \sgn (X_n - \theta_{n-1}).
\end{equation*}
We make one of two assumptions on the stepsizes $\gamma_n$, which
are relatively standard: we always have $\gamma_n$ non-increasing, and
\begin{subequations}
  For some $0 < \lambda \le 1$,
  \begin{align}
    \label{eqn:lazy-gamma}
    \frac{\gamma_n - \gamma_{n+1}}{\gamma_n^2}
    \to 0, & ~~~
    \sum_n \frac{\gamma_n^\frac{1 + \lambda}{2}}{\sqrt{n}} < \infty
    ~~ \mbox{or} \\
    \gamma_n = o(n^{-2/3}),
    & ~~~
    \sum_n \gamma_n = \infty.
    \label{eqn:stringent-gamma}
  \end{align}
\end{subequations}

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(0)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    Let $\{P_\theta\}_{\theta \in \R}$ be the family of distributions
    with density $f(\cdot - \theta)$, where $f$ has median 0.
    Let $h_n \to h \in \R$, and define the distributions
    $P_n = P_{\theta + h_n/\sqrt{n}}^n$. Then
    \begin{equation*}
      \sqrt{n}\left(\bar{\theta}_n - \theta - h_n / \sqrt{n}\right)
      \mathop{\cd}_{P_n}
      \normal\left(0, \frac{1}{4 f(0)^2}\right),
    \end{equation*}
    and for any bounded, symmetric, and quasi-convex function $L$,
    \begin{align} 
      & \sup_{c < \infty} \limsup_{n \to \infty}
      \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}
      \E_\tau \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[L(W/ 2 f(0)) \right],
        \label{eq:attaining_LAM}
    \end{align}
    where $W \sim \normal(0,1)$. 
    %
    %
  \item \label{item:sgd-ms-convergence} Assume the stepsizes $\gamma_n$
    satisfy both conditions~\eqref{eqn:lazy-gamma}
    and~\eqref{eqn:stringent-gamma}. Let
    $\pi$ be a distribution on $\R$. Then
    \begin{equation*}
      \int \E\Big[( \bar{\theta}_n - \theta )^2\Big] \pi(d\theta)
      = \frac{1}{4 n f(0)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}

@@

@@

  Now let
  $E_n = \E[B_i \mid T_n] = P(X_i \le T_n)$,
  so that $\var(B_i \mid T_n) = E_n(1 - E_n)$. Define also the random
  variable
  \begin{equation*}
    Y_n \defeq \sqrt{n}
    \frac{1}{\sqrt{E_n (1 - E_n)}}
    \bigg[\frac{1}{n}\sum_{i = 1}^n B_i - E_n\bigg],
  \end{equation*}
  and let $F_n(\cdot \mid T_n)$ be its cumulative distribution function.
  Then because
  \begin{equation*}
    \E\left[|B_i - E_n|^3 \mid T_n\right] \le E_n(1 - E_n)
    ~~ \mbox{we have} ~~
    \E\left[\frac{|B_i - E_n|^3}{(E_n(1 - E_n))^{3/2}} \mid T_n\right]
    \le \frac{1}{\sqrt{E_n(1 - E_n)}}.
  \end{equation*}
 The Berry-Esseen theorem implies that there exists a constant
  $C \le 1$ such that
  \begin{equation*}
    \sup_t \left|F_n(t \mid T_n) - \Phi(t) \right|
    \le \frac{C}{\sqrt{E_n (1 - E_n)} \sqrt{n}} \wedge 2,
  \end{equation*}
  where $\Phi$ is the standard Gaussian CDF.
  As $E_n (1 - E_n) \cas \frac{1}{4}$ by definition of the median,
  we have
  that (with probability 1)
  \begin{equation*}
    \sup_t \left|F_n(t \mid T_n) - \Phi(t)\right| \le \frac{C}{\sqrt{n}}
    ~~ \mbox{eventually}.
  \end{equation*}
  By dominated convergence and Jensen's inequality we thus obtain
  \begin{equation*}
    \sup_t \left|\mathbb{P}(Y_n \le t) - \Phi(t)\right|
    \le \E\left[
      \sup_t \left|F_n(t \mid T_n) - \Phi(t)\right| \right]
    \to 0,
  \end{equation*}
  which gives that $Y_n \cd \normal(0, 1)$.

### 2019-12-23 conclusions.tex
corrected various typos

\section{Conclusions \label{sec:conclusions}}
We considered the risk and efficiency in estimating the mean of a symmetric and log-concave distribution from a sequence of bits, where each bit is obtained by encoding a single sample from this distribution. 
%
In an adaptive encoding setting, we showed that no estimator can attain asymptotic relative efficiency larger than that of the median of the samples. We also showed that this bound is tight by presenting two adaptive encoding and estimation procedures that are as efficient as the median. 
%We also characterized the one-step optimal scheme in this adaptive setting, i.e., the scheme that minimizes the risk given any set of previously obtained bits.
\par
In the distributed setting we provided conditions for local asymptotic normality of the encoded samples, which implies asymptotic minimax bound on both the risk and efficiency relative to the mean. 
%
These condition implies that, in the distributed, the optimal estimation performance derived for the adaptive case can only be attained over a finite number of points, i.e., no scheme is uniformly optimal in this setting. 
%
We further considered the special case where the sequence of bits is obtained by comparing against a prescribed sequence of thresholds. We characterized the performance of the optimal estimator from such bit-sequence using the density of the thresholds and considered the density that minimizes the minimax risk.

>>>>

\section{Conclusions \label{sec:conclusions}}
We considered the risk and efficiency in estimating the mean of a symmetric and log-concave distribution from a sequence of bits, where each bit is obtained by encoding a single sample from this distribution. 
%
In an adaptive encoding setting, we showed that no estimator can attain asymptotic relative efficiency larger than that of the median of the samples. We also showed that this bound is tight by presenting two adaptive encoding and estimation procedures that are as efficient as the median. 
%We also characterized the one-step optimal scheme in this adaptive setting, i.e., the scheme that minimizes the risk given any set of previously obtained bits.
\par
In the distributed setting we provided conditions for local asymptotic normality of the encoded samples, which implies asymptotic minimax bound on both the risk and efficiency relative to the mean. 
%
These condition implies that, in the distributed setting, the optimal estimation performance derived for the adaptive case can only be attained over a finite number of points, i.e., no scheme is uniformly optimal in this setting. 
%
We further considered the special case where the sequence of bits is obtained by comparing against a prescribed sequence of thresholds. We characterized the performance of the optimal estimator from such bit-sequence using the density of the thresholds and considered the density that minimizes the minimax risk.

### 2019-12-23 distributed-estimation.tex
corrected various typos

\caption{\label{fig:minimax_support}
Optimal threshold density $\lambda^\star(dt)$ (blue) that maximizes the ARE for $f(x) = \Ncal(\theta,\sigma^2)$ and $\theta \in [-T,T]$.
%
The continuous curve (red) represents the reciprocal of the asymptotic risk at a fixed $\theta \in [-T,T]$ under the optimal density, i.e., the minimax risk is the inverse of the minimal value of this curve. The dashed curve (green) is the reciprocal of the asymptotic risk at a fixed $\theta \in [-T,T]$ when the threshold values are uniformly distributed over $[-T,T]$, hence its minimal value is the inverse of \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

When a prior $\pi$ on $\Theta$ is provided, one may be interested in the threshold density $\lambda$ minimizing the Bayes risk $R_n(\pi)$. The resulting minimization problem is 
\begin{align}
\label{eq:cvx_average}
\begin{split}
\mathrm{minimize} \quad & R_{\pi} =  \int \frac{\pi(d\theta)}{ \int \eta \left( t-\theta\right) \lambda(dt)}. \\ 
\mathrm{subject~to} \quad & \lambda(dt)\geq 0,\quad \int \lambda(dt) =1, 
\end{split}
\end{align}
which is convex in $\lambda$ since $x \rightarrow 1/x$, $x>0$, is convex. Figure~\ref{fig:opt_density}  illustrates the solution to \eqref{eq:cvx_average} for the case of where $f(x)$ is the normal distribution and $\pi$ is the uniform over $\Theta = [-T,T]$. %Figure~\ref{fig:dist_bound_uniform} illustrates the corresponding Bayes risk.

>>>>

\caption{\label{fig:minimax_support}
Optimal threshold density $\lambda^\star(dt)$ (blue) that maximizes the ARE for $f(x) = \Ncal(\theta,\sigma^2)$ and $\theta \in [-T,T]$.
%
The continuous curve (red) represents the reciprocal of the asymptotic risk for each $\theta \in [-T,T]$ under the optimal density, i.e., the minimax risk is the inverse of the minimal value of this curve. The dashed curve (green) is the reciprocal of the asymptotic risk for each $\theta \in [-T,T]$ when the threshold values are uniformly distributed over $[-T,T]$, hence its minimal value is the inverse of \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

When a prior $\pi$ on $\Theta$ is provided, one may be interested in the threshold density $\lambda$ minimizing the Bayes risk $R_n(\pi)$. The resulting minimization problem is 
\begin{align}
\label{eq:cvx_average}
\begin{split}
\mathrm{minimize} \quad & R_{\pi} =  \int \frac{\pi(d\theta)}{ \int \eta \left( t-\theta\right) \lambda(dt)}. \\ 
\mathrm{subject~to} \quad & \lambda(dt)\geq 0,\quad \int \lambda(dt) =1, 
\end{split}
\end{align}
which is convex in $\lambda$ since $x \rightarrow 1/x$, $x>0$, is convex. Figure~\ref{fig:opt_density}  illustrates the solution to \eqref{eq:cvx_average} for the case of where $f(x)$ is the normal distribution and $\pi$ is the uniform distribution over $\Theta = [-T,T]$. %Figure~\ref{fig:dist_bound_uniform} illustrates the corresponding Bayes risk.

### 2019-12-23 preliminaries.tex
corrected various typos

A related problem is the CEO problem, which considers the estimation of a
sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$
is available at $n$ terminals. At each terminal $i$, an encoder observes the
$k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator~\cite{berger1996ceo}. The
central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with
the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the
distributed encoding setting is given by the minimal MSE in the CEO setting
as $k \to \infty$. Note that the difference between the CEO setting and ours
lays in the privilege of each of the encoders to describe $k$ realizations
of $\theta$ using $k$ bits with MSE averaged over these realizations, rather
than a single realization using a single bit in ours.

When the prior on $\theta$ and the noise corrupting it at each location are
Gaussian, Prabhakaran et al.~\cite{prabhakaran2004rate} characterize the
optimal encoding and its asymptotic risk as $k \to \infty$.  Chen et
al.~\cite{chen2004upper} also provide an expression for the quadratic risk
in the CEO setting under Gaussian priors. Adapting to our setting, this
expression provides the following proposition:
\begin{prop} \label{prop:CEO}
  Assume that $\Theta = \mathbb R$ and $\pi(\theta) =
  \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is
  arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed
  setting satisfies
  \begin{equation} \label{eq:ceo_bound}
    n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
  \end{equation}
  where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}
\noindent
See Appendix~\ref{app:proof:CEO} for a proof.

>>>>

A related problem is the CEO problem, which considers the estimation of a
sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$
is available at $n$ terminals. At each terminal $i$, an encoder observes the
$k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator~\cite{berger1996ceo}. The
central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with
the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta} \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the
distributed encoding setting is given by the minimal MSE in the CEO setting
as $k \to \infty$. Note that the difference between the CEO setting and ours
lays in the privilege of each of the encoders to describe $k$ realizations
of $\theta$ using $k$ bits with MSE averaged over these realizations, rather
than a single realization using a single bit in ours.

When the prior on $\theta$ and the noise corrupting it at each location are
Gaussian, Prabhakaran et al.~\cite{prabhakaran2004rate} characterize the
optimal encoding and its asymptotic risk as $k \to \infty$.  Chen et
al.~\cite{chen2004upper} also provide an expression for the quadratic risk
in the CEO setting under Gaussian priors. Adapting to our setting, this
expression provides the following proposition:
\begin{prop} \label{prop:CEO}
  Assume that $\Theta = \mathbb R$ and $\pi(\theta) =
  \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is
  arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed
  setting satisfies
  \begin{equation} \label{eq:ceo_bound}
    n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
  \end{equation}
  where the expectation is with respect to $\theta$ and $X_1,\ldots,X_n$.
\end{prop}
\noindent
See Appendix~\ref{app:proof:CEO} for a proof.

### 2019-12-23 problem-formulation.tex
corrected various typos

For $f$ the normal density, classical results~\cite{Samford1953,
  hammersley1950estimating} show that that $\eta(x)$ is a strictly
decreasing function of $|x|$, as we illustrate in Fig.~\ref{fig:eta}.
%
We consider log-concave symmetric distributions sharing this
property.  Specifically, we require the following.
\begin{assumption}
  \label{assump:failure_rate}
  The density $f$ is log-concave and symmetric.  Additionally, the origin $x
  = 0$ uniquely maximizes $\eta(x)$, and $\eta(x)$ is non-increasing in
  $|x|$.
\end{assumption}
Under this assumption,
\begin{equation*}
  4 f^2(x) \leq \eta(x) \leq \eta(0),
\end{equation*} 
%
where $\eta(0) = 4 f^2(0)$ is the asymptotic variance of the sample median
(Eq.~\eqref{eqn:variance-quantiles} at $p = 1/2$).  Combined with
log-concavity of $f(x)$, Assumption~\ref{assump:failure_rate} implies that
$\eta(x)$ vanishes as $|x|\rightarrow \infty$.  Several distributions
satisfy Assumption~\ref{assump:failure_rate}, including the generalized
normal distributions with a shape parameter between $1$ and $2$ (including
the normal and Laplace distributions). Symmetric log-concave distributions
failing Assumption~\ref{assump:failure_rate} include the uniform
distribution and the generalized normal distribution with shape parameter
greater than $2$. Some restriction on the class of distributions is
necessary to develop our results; indeed, in
Appendix~\ref{sec:uniform-weirdos} we provide a brief discussion on the
uniform distribution, where a one-step adaptive estimator with single bit
observations can achieve convergence rates faster than the familiar
$\sqrt{n}$ paramateric rate.

>>>>

For $f$ the normal density, classical results~\cite{Samford1953,
  hammersley1950estimating} show that $\eta(x)$ is a strictly
decreasing function of $|x|$, as we illustrate in Fig.~\ref{fig:eta}.
%
We consider log-concave symmetric distributions sharing this
property.  Specifically, we require the following.
\begin{assumption}
  \label{assump:failure_rate}
  The density $f$ is log-concave and symmetric.  Additionally, the origin $x
  = 0$ uniquely maximizes $\eta(x)$, and $\eta(x)$ is non-increasing in
  $|x|$.
\end{assumption}
Under this assumption,
\begin{equation*}
  4 f^2(x) \leq \eta(x) \leq \eta(0),
\end{equation*} 
%
where $\eta(0) = 4 f^2(0)$ is the asymptotic variance of the sample median
(Eq.~\eqref{eqn:variance-quantiles} at $p = 1/2$).  Combined with
log-concavity of $f(x)$, Assumption~\ref{assump:failure_rate} implies that
$\eta(x)$ vanishes as $|x|\rightarrow \infty$.  Several distributions
satisfy Assumption~\ref{assump:failure_rate}, including the generalized
normal distributions with a shape parameter between $1$ and $2$ (including
the normal and Laplace distributions). Symmetric log-concave distributions
failing Assumption~\ref{assump:failure_rate} include the uniform
distribution and the generalized normal distribution with shape parameter
greater than $2$. Some restriction on the class of distributions is
necessary to develop our results; indeed, in
Appendix~\ref{sec:uniform-weirdos} we provide a brief discussion on the
uniform distribution, where a one-step adaptive estimator with single bit
observations can achieve convergence rates faster than the familiar
$\sqrt{n}$ paramateric rate.
